Cover Times of Random Walks

### 2019-04-29 paper.tex
abstract, outline and general intro

\abstract{\todo{abstract}}

@@

\todo{write general introduction to cover times}
"white screen" problem
numerous applications
many mathematical tools

\subsection{Motivation}
\todo{write about applications}
universal travel sequences
graph connectivity
protocol testing
birthday problem
coupon collectors

\subsection{Definitions}
We consider an undirected, connected graph $G$ without self loops or multi-edges.
Let $V(G)$ be the set of vertices and let $E(G)$ be the set of edges on $G$.
Then $n$ is the number of vertices $|V(G)|$.
We say $i\sim j$ for $i$ and $j$ in the vertex set $V(G)$ if $(i,j)$ in $E(G)$.
Assume that each vertex is accessible from every other vertex.
That is, there is a set of edges from any $x$ to any $y$.
%Without loss of generality, we label the vertices $\{1, 2, 3, ..., n\}$.

Let $C_{i,j}$ be the non-negative weight of the edge between vertices $i$ and $j$.
We will mostly consider unweighted graphs
with $C_{i,j} = 1$ if $i \sim j$ and $C_{i,j} = 0$ otherwise.

@@

We consider a random walk $(X_t)$ on $G$.
Call $X_t$ the vertex that the walk is on at time $t$.
The random walker begins at vertex $X_0$.
At each step, the walker at vertex $i$ moves to neighboring vertex
$j$ with probability $c(i,j)/c(i)$.

@@

We now introduce voltage, current, and effective resistance.
Consider $a$ and $b$ in $V(G)$.
Set the voltage to be 1 at $a$ and 0 at $b$.
We may think of voltage as the probability we get to $a$ before $b$.
Probabilistically, of course $v_a = 1$ and $v_b=0$.
Voltage is a harmonic function so the voltage at every vertex $x$ in 
$V(G) \textbackslash \{a,b\}$ is determined by the boundary values
(at $a$ and $b$).

@@

Rely on \cite{BH94}
Always possible to find cover time but it gets hard as the graph gets bigger
Structure helps:
Complete graph,
Line graph,
Simple cycle (extended line graph)

In lieu of structure, we want to work with arbitrary graphs
Rely on \cite{DS84} for commute time relationship
Rely on \cite{LP17}
Matthews method is intuitive approach that uses same ideas as complete graph
Matthews method applied to binary tree with commute time relationship
Use \cite{AF14}
Explore the idea of a spanning tree and use it to bound time with commute identity

So far we've only worked with expected time
\cite{Du11}
Explore what happens for the variable and it's distribution

\todo{outline of each section and subsections}

\section{Cover Times of Structured Graphs}

\subsection{Small Example}
In general, it is always possible to find the exact cover time
of a graph using first step analysis.
We demonstrate the strategy on a small graph suggested by \cite{BH94}.

@@

\subsection{Complete Graph \label{sec:complete}}
We call the graph with an edge between each pair of vertices the
complete graph.
Note that a complete graph with $n$ vertices has ${n \choose 2}$ edges.

@@

A natural interpretation of the cover time on a complete graph is 
the so-called coupon collecting problem.
There are $r$ different coupons that are randomly packaged in cereal boxes.
An avid fan buys a box of cereal every day.
We want to know how many days until the fan has collected each coupon.
The strategy is to think of every coupon as a vertex on the complete graph.
We then move from vertex to vertex with uniform probability.
The difference between the coupon collecting problem and the complete graph
is that we now have self-edges.
Before, we had to leave our current vertex at each step.
Now, we can stay in the same vertex (provided our collector found the same coupon
two days in a row).
We apply our approach to the complete graph and substitute the $r$ possible coupons
we can find for the $n-1$ vertices we could move to.
Then
\begin{align}
\tcov&=  r \left(\frac{1}{r}+\frac{1}{r-1}+\ldots+1\right). \nonumber
\end{align}

@@

\subsection{Line Graph}
A line graph is a set of vertices connected in a line.
(\cref{fig:line} is an example with $n=N+1$.)

@@

\subsection{Simple Cycle}

@@

\section{Bounds}

@@

\subsection{Matthews Method}
We have seen that it is in general very hard
to calculate the cover time of a random walk
on a graph.
The exceptions are symmetric or small graphs.
In lieu of an exact solution, we want to 
bound the cover time.

@@

\subsection{Matthews Method on Binary Trees}

@@

\subsection{Spanning Tree Argument}
\cite{AF14}
We introduce the concept of a spanning tree to create
another upper bound for the cover time of a random walk.

@@

\section{Cover Time Distributions}
\subsection{Complete Graph}
\cite{Du11}
We consider the distribution of the cover time of a random walk
on the complete graph with self edges.
Call $r$ the number of steps since we began the random walk.
So that the notation works out more easily,
set $r=1$ when we begin our walk at $X_0$.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
P(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

@@

\subsection{The Birthday Problem}
Consider a complete graph (with self-edges) and 365 vertices.
We may think of the cover time variable of this graph as the number of people
until every birthday is represented.
We call this the birthday problem.

@@

There appears to be a critical stage where it quickly becomes very likely
that all birthdays have been represented.
\cref{fig:birthday} shows the cumulative distribution of the birthday problem.

@@

\newpage
\begin{appendices}
\subsection{Birthday Problem}
\lstinputlisting{computation.R}
\end{appendices}

>>>>

\abstract{
The cover time of a graph is the time it takes for a random walk to
visit all of its vertices.
We provide a detailed overview for a variety of cover time topics:
We begin by investigating the cover time of small and structured graphs but
quickly need a more robust strategy for larger, unstructured ones.
We apply more general probablistic arguments and an important electrical
network relationship to identify bounds.
Finally, we describe the limiting distribution of the cover time of
complete graphs with self-edges and apply our analysis to an intuitive example.
}

@@

We begin by describing two intuitive examples of cover time problems.
Consider a coupon collector who opens a single box of cereal everyday.
In each box, is one of $n$ distinct coupons.
How many days until our collector has collected at least one of every coupon?
(The coupon collector's problem corresponds to the cover time of the complete
graph with self-edges.)

Next consider a similar scenario called the birthday problem.
How many people would we need to ensure that every day of the year
(excluding February 29) is represented as a birthday?
A lower bound is certainly 365 but it turns out that there is no upper bound.

Other applications of cover times include universal travel sequences,
graph connectivity algorithms, and protocol testing.

In this paper, we use a variety of strategies like first step analysis
and draw on several related topics such as electrical networks to rigorously
investigate cover times.

\subsection{Definitions}
We consider an undirected, connected graph $G$ without self- or multi-edges.
Let $V(G)$ be the set of vertices and let $E(G)$ be the set of edges on $G$.
Then $n$ is the number of vertices $|V(G)|$.
We say $i\sim j$ for $i$ and $j$ in the vertex set $V(G)$ if $(i,j)$ is in $E(G)$.
Assume that each vertex is accessible from every other vertex.
That is, there is a set of edges connecting any $x$ to any $y$.
%Without loss of generality, we label the vertices $\{1, 2, 3, ..., n\}$.

Let $C_{i,j}=C_{j,i}$ be the non-negative weight of the edge between vertices $i$ and $j$.
We will mostly consider unweighted graphs
such that $C_{i,j} = 1$ if $i \sim j$ and $C_{i,j} = 0$ otherwise.

@@

We consider a random walk $(X_t)$ on $G$.
Call $X_t$ the vertex that the walk is on at time $t$.
The random walk begins at vertex $X_0$.
At each step, the walk at vertex $i$ moves to neighboring vertex
$j$ with probability $C(i,j)/C(i)$.

@@

We now introduce voltage, current, and effective resistance.
Consider $a$ and $b$ in $V(G)$.
Set the voltage to be 1 at $a$ and 0 at $b$.
We may think of voltage as the probability we get to $a$ before $b$.
Probabilistically, of course $v_a = 1$ and $v_b=0$.
The voltage at every vertex $x$ in  $V(G) \textbackslash \{a,b\}$
is determined by the boundary values (at $a$ and $b$).

@@

We begin in \cref{sec:small_example}.
(\cref{sec:structured_graphs} primarily relies on \cite{BH94}.)
We see that it is always possible to exactly calculate the cover time
of graph but it is in general very difficult to do so.
For the remainder of this section, we investigate
structured graphs that are exceptions to the rule:
the complete graph, line graphs, and simple cycles.

\cref{sec:bounds} deals with the large or unstructured graphs 
for which we must use bounds.
We begin by establishing an important identity between
commute times and effective resistance in \cref{sec:commute}.
(Our electrical network analysis is an extension of \cite{DS84}
and we use \cite{LP17} for the next two subsections.)
We take our strategy for finding the cover time of the complete graph
from \cref{sec:complete} and apply it to general graphs in \cref{sec:matthews}.
In \cref{sec:binary}, we use both the commute time identity and the Matthews Method
to bound the cover time of a binary tree.
In \cref{sec:spanning}, we follow \cite{AF14} and use the idea of a spanning tree
as well as the commute time identity to develop another bound for general graphs.

Up until \cref{sec:dist}, we have only considered the expected cover time.
We now delve into the distribution of the cover time of a complete graph.
We draw on \cite{Du11} and find that the cover time variable approaches a Poisson
distribution.
In \cref{sec:birthday}, we apply \cref{sec:complete_dist} to the birthday problem.
(The R code for our computations and \cref{fig:birthday} can be found in \cref{app:birthday}.)

\section{Cover Times of Structured Graphs}\label{sec:structured_graphs}

\subsection{Small Example}\label{sec:small_example}
In general, it is always possible to find the exact cover time
of a graph using first step analysis.
We demonstrate the strategy on a small graph suggested by \cite{BH94}.

@@

\subsection{Complete Graph}\label{sec:complete}
We call the graph with an edge between each pair of vertices the
complete graph.
Note that a complete graph with $n$ vertices has ${n \choose 2}$ edges.

@@

A natural interpretation of the cover time on a complete graph is 
the so-called coupon collecting problem.
There are $r$ different coupons that are randomly packaged in cereal boxes.
An avid fan buys a box of cereal every day.
We want to know how many days until the fan has collected each coupon.
The strategy is to think of every coupon as a vertex on the complete graph.
We then move from vertex to vertex with uniform probability.
The difference between the coupon collecting problem and the complete graph
is that we have self-edges
whereas before we had to leave our current vertex at each step.
Now, we can stay in the same vertex (provided our collector found the same coupon
two days in a row).
We apply our approach to the complete graph and substitute the $r$ possible coupons
we can find for the $n-1$ vertices we could move to.
Then
\begin{align}
\tcov&=  r \left(\frac{1}{r}+\frac{1}{r-1}+\ldots+1\right). \nonumber
\end{align}

@@

\subsection{Line Graph}\label{sec:line_graph}
A line graph is a set of vertices connected in a line.
(\cref{fig:line} is an example with $n=N+1$.)

@@

\subsection{Simple Cycle}\label{sec:simple_cycle}

@@

\section{Bounds}\label{sec:bounds}

@@

\subsection{Matthews Method}\label{sec:matthews}
We have seen that it is in general very hard
to calculate the cover time of a random walk
on a graph.
The exceptions are symmetric or small graphs.
In lieu of an exact solution, we want to 
bound the cover time.

@@

\subsection{Matthews Method on Binary Trees}\label{sec:binary}

@@

\subsection{Spanning Tree Argument}\label{sec:spanning}
\cite{AF14}
We introduce the concept of a spanning tree to create
another upper bound for the cover time of a random walk.

@@

\section{Cover Time Distributions}\label{sec:dist}
\subsection{Complete Graph}\label{sec:complete_dist}
\cite{Du11}
We consider the distribution of the cover time of a random walk
on the complete graph with self edges.
Call $r$ the number of steps since we began the random walk.
So that the notation works out more easily,
set $r=1$ when we begin our walk at $X_0$.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
P(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

@@

\subsection{The Birthday Problem}\label{sec:birthday}
Consider a complete graph (with self-edges) and 365 vertices.
We may think of the cover time variable of this graph as the number of people
until every birthday is represented.
We call this the birthday problem.

@@

There appears to be a critical threshold where it quickly becomes very likely
that all birthdays have been represented.
\cref{fig:birthday} verifies our intuition.

@@

\newpage
\begin{appendices}
\subsection{Birthday Problem}\label{app:birthday}
\lstinputlisting{computation.R}
\end{appendices}
